### 2018-07-11 paper/supplementary/supplementary.tex
Add supplementary information .tex file and auxiliaries

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when producing your supplementary material for your Frontiers article.                                 %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 2.5 Generated 2018/06/15 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
%%%  NB logo1.jpg is required in the path in order to correctly compile front page header %%%

\documentclass[utf8]{frontiers_suppmat} % for all articles
\usepackage{url,hyperref,lineno,microtype}
\usepackage[onehalfspacing]{setspace}

% bibliography
\usepackage[english]{babel}
\usepackage{csquotes}
\usepackage[
backend=biber, 
bibencoding=utf8, 
style=authoryear, 
date=year, 
maxbibnames=6, 
minbibnames=6, 
url=false, 
isbn=false, 
eprint=false,
uniquename = false, 
autocite=inline]{biblatex}
\addbibresource{conductances.bib}

% make pretty MATLAB code
\usepackage{listings}
\usepackage{matlab-prettifier}
\lstMakeShortInline[style=Matlab-editor]"

% Leave a blank line between paragraphs instead of using \\

\begin{document}
\onecolumn
\firstpage{1}

\title[Supplementary Material]{{\helveticaitalic{Supplementary Material}}:
\\ \helvetica{Xolotl: An Intuitive and Approachable Neuron \& Network Simulator in MATLAB}} %Please insert the title of your article here

\maketitle

\section{Database of Conductances}

Dynamics for model conductances in the following papers have been implemented in "xolotl".

\begin{enumerate}
	\item \cite{bronkRegulationEagCa22018}
	\item \cite{brookingsAutomaticParameterEstimation2014}
	\item \cite{caplanManyParameterSets2014a}
	\item \cite{dethierPositiveFeedbackCellular2015a}
	\item \cite{drionDopaminePacemakerNeuron2017}
	\item \cite{goldmanGlobalStructureRobustness2001a}
	\item \cite{herasModulationVoltagedependentConductances2018}
	\item \cite{kisperskyIncreaseSodiumConductance2012a}
	\item \cite{linActivityDependentAlternativeSplicing2012}
	\item \cite{liuModelNeuronActivityDependent1998}
	\item \cite{nadimFrequencyRegulationSlow1998}
	\item \cite{nadimSynapticDepressionCreates1999}
	\item \cite{odowdVoltageclampAnalysisSodium1988}
	\item \cite{prinzAlternativeHandtuningConductancebased2003a}
	\item \cite{rodriguezConvergentRhythmGeneration2013}
	\item \cite{sharpDynamicClampComputergenerated1993a}
	\item \cite{soto-trevinoActivitydependentModificationInhibitory2001}
	\item \cite{soto-trevinoComputationalModelElectrically2005a}
	\item \cite{swensenModulatorsConvergentCellular2001a}
	\item \cite{turrigianoSelectiveRegulationCurrent1995a}
	\item \cite{wicherNonsynapticIonChannels2001}
\end{enumerate}

\printbibliography

\end{document}