### 2018-07-11 paper/supplementary/supplementary.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% This is just an example/guide for you to refer to when producing your supplementary material for your Frontiers article.                                 %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Version 2.5 Generated 2018/06/15 %%%
%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
%%%  NB logo1.jpg is required in the path in order to correctly compile front page header %%%

\documentclass[utf8]{frontiers_suppmat} % for all articles
\usepackage{url,hyperref,lineno,microtype}
\usepackage[onehalfspacing]{setspace}

% bibliography
\usepackage[english]{babel}
\usepackage{csquotes}
\usepackage[
backend=biber, 
bibencoding=utf8, 
style=authoryear, 
date=year, 
maxbibnames=6, 
minbibnames=6, 
url=false, 
isbn=false, 
eprint=false,
uniquename = false, 
autocite=inline]{biblatex}
\addbibresource{conductances.bib}

% make pretty MATLAB code
\usepackage{listings}
\usepackage{matlab-prettifier}
\lstMakeShortInline[style=Matlab-editor]"

% Leave a blank line between paragraphs instead of using \\

\begin{document}
\onecolumn
\firstpage{1}

\title[Supplementary Material]{{\helveticaitalic{Supplementary Material}}:
\\ \helvetica{Xolotl: An Intuitive and Approachable Neuron \& Network Simulator in MATLAB}} %Please insert the title of your article here

\maketitle

\section{Database of Conductances}

Dynamics for model conductances in the following papers have been implemented in "xolotl".

\begin{enumerate}
	\item EAGes, EAGmut, EAGwt (\cite{bronkRegulationEagCa22018})
	\item A-type (\cite{brookingsAutomaticParameterEstimation2014})
	\item MICurrent (\cite{caplanManyParameterSets2014a})
	\item CaT, H, Kd, NaV (\cite{dethierPositiveFeedbackCellular2015a})
	\item CaL, CaPump, KCa, Kd, NaV (\cite{drionDopaminePacemakerNeuron2017})
	\item Ap, At, Cal, KCa, KCaT, Proc (\cite{goldmanGlobalStructureRobustness2001a})
	\item Shaker (\cite{hardieNovelPotassiumChannels1991})
	\item Novel, Shab, Shaker (\cite{herasModulationVoltagedependentConductances2018})
	\item A-type, CaS, CaT, H, KCa, Kd, NaV (\cite{kisperskyIncreaseSodiumConductance2012a})
	\item NaP, NaT (\cite{linActivityDependentAlternativeSplicing2012})
	\item A-type, CaS, CaT, H, KCa, Kd, NaV (\cite{liuModelNeuronActivityDependent1998})
	\item \cite{liuModelNeuronActivityDependent1998} conductances with cached gating functions
	\item \cite{liuModelNeuronActivityDependent1998} conductances with forward Euler integration
	\item \cite{liuModelNeuronActivityDependent1998} conductances with temperature dependence
	\item Kd and NaV conductances from Int1, LG, and MCN1 cells (\cite{nadimFrequencyRegulationSlow1998})
	\item Cal (\cite{nadimSynapticDepressionCreates1999})
	\item Drosophila NaV (\cite{odowdVoltageclampAnalysisSodium1988})
	\item A-type, CaS, CaT, H, KCa, Kd, NaV (\cite{prinzAlternativeHandtuningConductancebased2003a})
	\item \cite{prinzAlternativeHandtuningConductancebased2003a} conductances with cached gating functions
	\item \cite{prinzAlternativeHandtuningConductancebased2003a} conductances with temperature dependence
	\item Kd and NaV from Int1 cells and modulatory input and transient LTS conductances from LG cells (\cite{rodriguezConvergentRhythmGeneration2013})
	\item proctolin modulatory input (\cite{sharpDynamicClampComputergenerated1993a})
	\item Cal and Kd from AB-PD, LP, and PY cells (\cite{soto-trevinoActivitydependentModificationInhibitory2001})
	\item A-type, CaT, and KCa from AB and PD cells, generic H, Kd, NaP, NaV, and proctolin conductances (\cite{soto-trevinoComputationalModelElectrically2005a})
	\item proctolin modulatory input (\cite{swensenModulatorsConvergentCellular2001a})
	\item A-type, CaS, CaT, H, KCa, Kd, NaP, NaV (\cite{turrigianoSelectiveRegulationCurrent1995a})
	\item Drosophila NaV (\cite{wicherNonsynapticIonChannels2001})
\end{enumerate}

\printbibliography

\end{document}