### 2018-08-14 paper/paper.tex
getting close to the final stretch

% !BIB TS-program = biber
\documentclass{frontiersSCNS} % for Science articles

\usepackage{graphicx}
%
\usepackage{amsmath}
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
\usepackage{color, soul}

\usepackage{caption}
\usepackage{float}
\floatstyle{plaintop}

\usepackage{hyperref}

% bibliography
\usepackage[english]{babel}
\usepackage{csquotes}
\usepackage[
backend=biber,
bibencoding=utf8,
style=authoryear,
date=year,
maxbibnames=6,
minbibnames=6,
url=false,
isbn=false,
eprint=false,
uniquename = false,
autocite=inline]{biblatex}
\addbibresource{Bibliography.bib}

\usepackage[draft, nomargin, marginclue, footnote]{fixme}
\fxsetup{targetlayout=color}

% new commands and shortcuts
\usepackage{xspace}
\newcommand{\eg}{\textit{e.g.}\xspace}
\newcommand{\etal}{\textit{et al.}\xspace}
\newcommand{\ie}{\textit{i.e.}\xspace}
\newcommand{\etc}{\textit{etc.}\xspace}
\newcommand{\vs}{\textit{vs.}\xspace}
\newcommand{\viz}{\textit{viz.}\xspace}

% shortcut for float placing
\usepackage{placeins}

% make pretty MATLAB code
\usepackage{listings}
\usepackage{matlab-prettifier}
\lstMakeShortInline[style=Matlab-editor]"

% other nice features
\usepackage{lineno}
\linenumbers

\usepackage{xifthen}% provides \isempty test

\copyrightyear{}
\pubyear{}

\def\journal{Neuroinformatics}
\def\DOI{}
\def\articleType{Technology Report}
\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Gorur-Shandilya, Hoyland \& Marder}
\def\Authors{Srinivas Gorur-Shandilya\,$^{1*\dagger}$, Alec Hoyland\,$^{1\dagger}$, and Eve Marder$^{1}$}
\def\Address{$^{1}$Volen National Center for Complex Systems and Biology Department \\ Brandeis University \\ Waltham, MA 02453 \\ USA
\vspace{0.25cm}

\vspace{0.5em}\tiny{$\dagger$ These authors have equally contributed to this article.}
}
\def\corrAuthor{Srinivas Gorur-Shandilya }
%\def\corrAddress{Marder Lab, Brandeis University, Biology Department and Volen Center for Complex Systems, Waltham, MA, USA}
\def\corrAddress{Volen National Center for Complex Systems, Brandeis University, Waltham MA 02453}
\def\corrEmail{\texttt{srinivasgs@brandeis.edu}}

\newcommand{\dd}[1]{\mathrm{d}#1}

\begin{document}
\onecolumn
\firstpage{1}

\title[xolotl: neuron \& network simulator]{xolotl: An Intuitive and Approachable Neuron \& Network Simulator}
\author[\firstAuthorLast ]{\Authors}
\address{}
\correspondance{}
\extraAuth{}
\topic{An Intuitive Neuronal Simulator}

\maketitle

\begin{abstract}

	Conductance-based models of neurons are used extensively in computational neuroscience, and working with these models continues to be a challenge due to their high dimensionality and large number of parameters. Here, we present a neuron and network simulator built on a novel automatic type system that binds object-oriented code written in C++ to objects in MATLAB. Our approach builds on the tradition of uniting the speed and power of languages like C++ with the ease-of-use and feature-set of scientific programming languages like MATLAB. Our software allows for the creation and manipulation of hierarchical models with components that are named and searchable, permitting intuitive high-level programmatic control over all parameters. Our simulator's architecture allows for the live manipulation of any parameter in any model, and for visualizing the effects of changing that parameter immediately. Our simulator is fully featured with hundreds of ion channel models from the electrophysiological literature, and can be extended to include arbitrary conductances, synapses, and mechanisms. Several core features like bookmarking of parameters and automatic hashing of source code promote reproducible and auditable research. Finally, our simulator is written in a modular fashion and has been released under a permissive free software license, enabling seamless integration into the workflow of other researchers.

	\tiny
	\keyFont{ \section{Keywords:} simulator, code:MATLAB, code:C++, conductance-based, neuron, network, neuron models}

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  								%
%		INTRODUCTION 			%
% 		& OVERVIEW   			%
%								%
%								%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:intro}

Nervous systems process and transmit information using electrically excitable membranes. Conductance-based models are a powerful biophysical simplification of an electrically excitable compartment in a neuron (\cite{hodgkinQuantitativeDescriptionMembrane1952}). Studies based on the Hodgkin-Huxley formalism now contribute significantly to mainstream research in small circuits (\cite{marderTheoryMotion1995, prinzComputationalApproachesNeuronal2010, prinzInsightsModelsRhythmic2006}). These models provide an approachable framework for understanding many salient principles of electrophysiology, since they explicitly model cell membranes and ion channels as electrical components in a circuit. However, several challenges remain in working with biophysically-detailed conductance-based neuron models. First, these models can be high-dimensional with many nonlinear differential equations, each with several parameters. Second, many or all equations in these models can be strongly coupled through dynamical variables like the membrane potential.  In multi-compartment models of spatially extended neurons, membrane potentials in every compartment can be different, and are coupled to each other. Finally, the choice of programming language used to implement the model imposes tradeoffs in designing and using neuron and network simulators: simulators written in languages like C++ or FORTRAN can integrate equations quickly, but often lack the ease-of-use and interoperability of those written in scientific programming languages like Python, Julia, or MATLAB (Mathworks).

Two major approaches have dominated the design of neuron simulators. One approach is to write the simulator in a fast compiled language like C and allow for the construction and simulation of neuron models using object-oriented paradigms. This approach has been implemented in NEURON (\cite{hinesNEURONSimulationEnvironment1997}). Simulators designed in this way tend to perform fast computations with little overhead, but suffer from a steep learning curve. Wrapping these simulators in a more approachable language like "Python"  or using graphical user interfaces (GUIs) mitigates these drawbacks (\cite{hinesNEURONPython2009}) at the cost of obfuscating the underlying algorithms and parameters (\cite{bretteSimulationNetworksSpiking2007, hinesNEURONPython2009}). In contrast, simulators designed from the ground up in popular scientific computing languages can be easier to use and benefit from interoperability with other commonly-used tools. Simulators like DynaSim (\cite{sherfeyDynaSimMATLABToolbox2018}), "ANNarchy" (\cite{vitayANNarchyCodeGeneration2015}), "BRIAN" (\cite{stimbergBrianSecondComing2013}), "morphforge" (\cite{hullMorphforgeToolboxSimulating2014}), and "PyNN" (\cite{davisonPyNNCommonInterface2009}) allow the user to specify models with strings of equations or components that are constructed using a special syntax, that can then be translated into a faster implementation language such as "C" or C++ (\cite{stimbergEquationorientedSpecificationNeural2014}). This approach permits considerable flexibility for simulating systems of differential equations. Since models need to be translated between the two languages, the hierarchical nature of neuron models is not naturally encapsulated by these tools, and the syntax can be verbose. Neither approach facilitates the creation of tools that simultaneously maintain efficiency, ease-of-use, and clarity.

To overcome these design limitations, we have developed a novel automatic type system, that we call "cpplab", which binds MATLAB code to classes specified in C++ header files. This architecture automatically creates objects in MATLAB that represent the underlying object-oriented C++ code, allowing the symbolic manipulation of C++ objects in the MATLAB interface. In this paper, we introduce xolotl, an implementation of the "cpplab" system specialized in integrating conductance-based neuron and network models. Models can be easily constructed from network building blocks in a few lines of MATLAB code using a hierarchical and intuitive syntax. Since models in the MATLAB workspace are automatically linked to models in the C++ implementation, configuring these objects in MATLAB transparently configures the underlying C++ objects. xolotl comes packaged with hundreds of components that can be used to assemble cells and networks;  has built-in visualization functions to inspect voltage time traces and activation functions; and a GUI for real-time manipulation of model parameters. The ease-of-use of the tools built into our software lends them to pedagogical applications and the rapid prototyping of models. Our software aims to simplify the investigation of the dynamics of conductance-based network and neuron models, facilitate collaborative modeling, and is intended to complement other tools being developed in the computational neuroscience community.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		DESIGN GOALS
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Design Goals}
\label{design}

xolotl is designed to be easy-to-use and richly featured while being fast enough to use in everyday research. Our focus was on designing an approachable simulator of conductance-based neurons and networks of these neurons; simulating arbitrary dynamical systems is therefore beyond the scope of this software. Specifically, the software was designed to simulate models of the form

\begin{equation}
C_{i}\frac{dV_{i}}{dt}=-\sum_{j}I_{j} \label{eq:1}
\end{equation}

where \( C_{i} \) and  \( V_{i} \) are the capacitance and membrane potential of compartment  \( i \). Compartments can represent whole neurons or parts of neurons.  \( I_{j} \) is the current due to ion channel population \( j \) and is given by

\begin{equation}
I_{j}=\bar{g}_{j}m_{j}^{p}h_{j}^{q}(V-E_{j})A_{i}  \label{eq:2}
\end{equation}

Here, \( \bar{g}_{j} \) is the maximal conductance, and \( E_{j} \) is the reversal potential of the ion channel population \( j \). \( A_{i} \) is the surface area of compartment \( i\) that contains these ion channels.  \( m_{j} \) and \( h_{j} \) are activation and inactivation variables that change according to

\begin{eqnarray*}
\tau_{m}\frac{dm}{dt}=m_{\infty}-m & \mathrm{and} & \tau_{h}\frac{dh}{dt}=h_{\infty}-h
\end{eqnarray*}

Typically, \( \tau_{m} \), \( \tau_{h} \), \( m_{\infty} \), and \( h_{\infty} \) are functions of the membrane potential \( V_{i} \). The software uses integration schemes that have been specifically developed to solve equations of this form (\cite{dayanTheoreticalNeuroscience2001, hinesEfficientComputationBranched1984, ohErrorAnalysisSpecialized2006}), though other schemes can be used if desired.

This software is designed to be used from within MATLAB, a scientific programming language common amongst neuroscientists and engineers for pedagogy and research. Our goal was to make xolotl completely usable entirely from within MATLAB. Models created using this simulator appear in the MATLAB workspace as native objects, are are thus fully scriptable, and are fully compatible with the large library of toolboxes that MATLAB provides, allowing the software to be used as a component of other packages and tools. All parameters of models and activation functions of any channel can be inspected at any point. We designed several features of the simulator to be easily extensible: adding custom conductances or synapse types is possible by calling functions that generate new C++ files on-the-fly. Finally, xolotl fully auditable by design, with several tools to verify model and parameter integrity and aid in reproducibility.

\subsection{Features}
\label{features}

\paragraph{A rich library of network components.} xolotl comes packaged with hundreds of pre-existing components (compartments, synapses, conductances, and mechanisms) that can be used as building blocks to construct model neurons and networks. Compartments represent sections of membrane with a single membrane potential, Calcium level, and set of constituent components, representing either entire cells or parts of cells. Objects of type "conductance" represent populations of ion channels in a compartment that produce transmembrane currents. "synapse" objects connect two compartments together by introducing a current in the postsynaptic compartment that depends on the presynaptic compartment's membrane potential. Objects of type "mechanism" can represent any intracellular mechanism and can read and modify any other component in the cell, and can run arbitrary dynamical systems within them.

\paragraph{Object-oriented.}
xolotl is designed to mirror the nested and hierarchical structure of networks and neurons. Biological neuronal networks are made up of neurons that are connected to each other with synapses. Each of these neurons contains within it a set of conductances and synaptic currents that contribute to its electrical behavior. Intracellular mechanisms can act within the cell, or parts of the cell, to modify and regulate conductances, synapses, or other dynamic properties of the cell. Similarly, a xolotl model can contain a set of compartments that can represent individual neurons. Each compartment can contain an arbitrary set of conductances. Compartments, conductances and synapses can contain mechanisms that can affect anything in the model. All objects in the xolotl model tree (compartments, conductances, etc.) are bonafide MATLAB objects with their own type and properties. This object-oriented programming paradigm naturally represents the hierarchical structure of biological networks that contain neurons that contain populations of ion channels, and makes constructing, integrating, and thinking about these models easier.

\paragraph{Automatic type system.}
To circumvent the tradeoff between high-performance but hard to use languages like C++ and richly-featured but potentially slow languages like MATLAB, we constructed an automatic type system that links object oriented code in C++ to object oriented code in MATLAB. This architecture lets us construct the core of the simulator in C++, leveraging features of C++ like pointers that are not readily available in MATLAB. A rudimentary way to make this C++ code useable in MATLAB would be to re-write that code in MATLAB so that MATLAB objects can be bound to their C++ implementations. However, this approach is cumbersome and inefficient, and can introduce errors. Instead, our automatic type system creates objects in MATLAB on-the-fly from C++ class specifications, obviating the need to rewrite code in MATLAB while preserving a tight coupling between objects in the MATLAB workspace and their C++ implementation. Crucially, this method makes developing new code much easier and simplifies the task of constructing complex frameworks that span these two languages.

\paragraph{Automatic hashing and compiling.}
Since every model requires a compiled binary to run, a potential stumbling block is the problem of unambiguously linking a model to a binary executable. xolotl solves this problem by hashing (\cite{rivestMD5MessageDigestAlgorithm1992}) the C++ header files of every component in the model recursively, allowing a model, no matter how complex, to be compactly represented by a short alphanumeric identifier (its ``hash''). Compiled binaries are named using this hash, ensuring both that the correct binary is run to integrate the model, and that compilation occurs only as needed. This powerful feature enables the user experience to remain entirely within the MATLAB workspace, with compilation and selection of the correct binary occurring silently in the background.

\subsection{Limitations}
\label{limitations}

Our focus on our simulator's ease-of-use and speed imposed some limitations on its feature set.

\paragraph{Limited to conductance-based models.} xolotl has been developed specifically for conductance-based models. It does not currently support rate- or current-based models, or arbitrary dynamical systems.

\paragraph{Limited numerical integration strategies.} Most components in the software are integrated using the exponential Euler method, which has been shown to perform well in neuronal models (\cite{ohErrorAnalysisSpecialized2006, dayanTheoreticalNeuroscience2001}). However, it may be desirable to use other methods under certain conditions. It is possible to introduce new network components that implement other integration schemes, or to modify the integration schemes of existing components, but that requires writing new C++  code. Currently, xolotl can only implement integration schemes with fixed step size.

\paragraph{Inefficient tools for handling large networks.} xolotl was designed to work with small but complex networks and models, where every compartment and component is named, rather than numbered. It is more therefore suited towards simulating small and heterogeneous networks rather than large but homogenous networks. While the software can integrate large networks ($>1000$ compartments), other tools are presumably more suited to this task, offering more natural frameworks for dealing with a large number of identical units. Similarly, xolotl is not optimized to solve coupled ODEs on complex branched morphologies, and other simulators like (\cite{hinesNEURONSimulationEnvironment1997}) are more suited for this task.

\paragraph{New mechanisms require new \texttt{C++} code.} Adding new network components requires writing new C++ code. A new conductance in the Hodgkin-Huxley formalism (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952, hodgkinQuantitativeDescriptionMembrane1952, dayanTheoreticalNeuroscience2001}) requires creating a new C++ header file, though this is generally trivial. Implementing a new integration scheme or component type requires much more in-depth knowledge of the underlying C++ core code.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		USAGE EXAMPLES
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Usage Examples}
\label{usage}

In this section, we illustrate how xolotl can be used to generate, inspect, and simulate a variety of models. These examples have been chosen to illustrate various features of xolotl, and are intended to serve as templates upon which researchers and educators can build.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		SIMULATING A HODGKIN-HUXLEY MODEL
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Simulating a Hodgkin-Huxley Model}

The axon of the giant squid contains contains a fast inactivating sodium conductance ("NaV"), a slower non-inactivating potassium conductance ("Kd"), and a passive leak current ("Leak"). Seminal work by Hodgkin and Huxley showed that depolarizations of the membrane could lead to an activation of the voltage-sensitive "NaV" channels, which led to a run away depolarization that was terminated by the inactivation of "NaV" channels and the activation of "Kd" channels that repolarized the membrane (\cite{hodgkinComponentsMembraneConductance1952, hodgkinMeasurementCurrentvoltageRelations1952}). As one of the simplest models of excitable neural membranes, the Hodgkin-Huxley model often serves as a the first model introduced in pedagogical literature (\cite{dayanTheoreticalNeuroscience2001, sterrattPrinciplesComputationalModelling2011, trappenbergFundamentalsComputationalNeuroscience2010}).

In this example, we demonstrate how to simulate the spiking activity of a Hodgkin-Huxley-like model, and how the tools built into our simulator make it easy to set up and integrate the model and gain insight into the underlying biophysical mechanisms. This simple model consists of a single electrical compartment with three types of conductances (Figure \ref{fig:figurehh}A). This hierarchical organization of the neuron is mirrored in the structure of the model in our simulator: an object of type "compartment" is used to represent the cell body, and three objects of type "conductance" are used to model the three populations of ion channels (Figure \ref{fig:figurehh}B). These models of ion channels were obtained from \cite{liuModelNeuronActivityDependent1998} based on electrophysiological recordings from the lobster stomatogastric ganglion (\cite{turrigianoSelectiveRegulationCurrent1995}), and are part of the simulator. The code to set up this model is terse, idiomatic, relies on no special markup, and preserves the heirarchical nature of the model (Figure \ref{fig:figurehh}C).

Adding an injected current and calling the built-in "plot" function plots the time series of membrane voltage. In the absence of injected current, the model is quiescent. When 0.2 nA is injected, the model tonically spikes (Figure \ref{fig:figurehh}A-B).  The \texttt{plot} function displays a voltage trace colored by the dominant current (Figure \ref{fig:figurehh}A). Colors in the voltage trace indicate the strongest instantaneous inward current when the voltage is increasing, and strongest instantaneous outward current when the voltage is decreasing. This built-in feature reveals that the dominant current during the upswing of every action potential is the sodium current, and the dominant current immediately after the peak of the action potential is the potassium current, but that the leak current contributes to depolarization following an action potential. This feature could be useful in quickly understanding the contributions of a number of ion channel types in a complex voltage trace from a more complicated neuron model.

Integrating the model returns the voltage time series for every compartment:

\begin{lstlisting}[style=Matlab-editor]
V = x.integrate;
\end{lstlisting}

The model can be integrated for various amplitudes of injected current to determine its f-I curve, or the relationship between injected current and its firing rate (\cite{kisperskyIncreaseSodiumConductance2012}). Figure \ref{fig:figurehh}E shows the f-I curve of this model, obtained by repeated integration of the model.  Finally, the built-in \texttt{show} function can plot activation ($m$) and inactivation ($h$) curves and voltage-dependent timescales of any channel type in the simulator (Figure \ref{fig:figurehh}D-G). These plots reveal that activation kinetics of the "NaV" channels are much faster than that of the "Kd" channels (Figure \ref{fig:figurehh}F), which facilitates the transient depolarization in an action potential. In summary, the simulator allows the user to construct and integrate this model in a few lines of code, and provides rich visualization of the dynamics of the model.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		VOLTAGE CLAMP
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Performing a Voltage Clamp Experiment \textit{in-silico}}

Voltage clamping is an experimental technique where a amplifier is configured to inject the appropriate amount of current through a electrode to maintain the voltage of a cell at a desired value (\cite{dayanTheoreticalNeuroscience2001}). Under this paradigm, the membrane voltage is "clamped" or fixed to a desired value, permitting the study of voltage-dependent ion channels, since the sum of all currents through the population of ion channels in the cell is equal and opposite to the current injected by the clamp (Figure \ref{fig:figureclamp}A). By combining voltage clamp with the use of pharmacological agents to block all channels but the one of interest, the voltage-sensitivity of an ion channel population can be characterized (\cite{coleIonicCurrentMeasurements1960, coleIonsPotentialsNerve1955, hodgkinEffectSodiumIons1949, hodgkinMeasurementCurrentvoltageRelations1952, hodgkinQuantitativeDescriptionMembrane1952, turrigianoSelectiveRegulationCurrent1995}).

xolotl can reproduce such a voltage clamp experiment \textit{in-silico}. Figure \ref{fig:figureclamp}B illustrates how a simple model with a single compartment and a single ion channel type can be set up and clamped to a desired voltage.  Integrating the model yields the current required to clamp the cell at that voltage. Here, we use a delayed-rectifier potassium conductance (\cite{liuModelNeuronActivityDependent1998}) and simulate a voltage-clamp experiment whose goal is to infer the activation function of this channel. First, the cell is clamped to a number of different voltages (Fig. \ref{fig:figureclamp}C) and the resultant clamping currents are measured by integrating the model (Fig. \ref{fig:figureclamp}D). Since the compartment is being voltage clamped, integrating the model returns the clamping current:

\begin{lstlisting}[style=Matlab-editor]
I_clamp = x.integrate;
\end{lstlisting}

By repeating the integration at a number of clamping voltages, we observe that the asymptotic clamping currents depend on the clamping voltage in a nonlinear manner (Fig. \ref{fig:figureclamp}E), since the open probabilities of the channel are functions of the membrane voltage. Assuming the reversal potential is known, Eq. \eqref{eq:2} can be used to solve for the total conductance of the channel as a function of the clamping voltage (Fig. \ref{fig:figureclamp}F). Finally, a sigmoid can be fit to the normalized conductance-voltage curves to obtain the activation function of the ion channel population (Fig. \ref{fig:figureclamp}G-H). xolotl can therefore be used to describe graphically the theoretical underpinnings of ion channel characterization through voltage clamp and can serve as an effective pedagogical tool in computational neuroscience.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		MECHANISMS
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Intracellular mechanisms}

So far, the models we described only considered the voltage dynamics of a cell (the solution to Eq. \ref{eq:1}). However, real neurons possess several dynamical features, arising from a variety of intracellular mechanisms. xolotl makes it possible to model and include arbitrary intracellular mechanisms. In our simulator, these intracellular mechanisms are represented by the "mechanism" object, and can be bound to compartments, conductances, and other object types.

A key intracellular mechanisms is the cytosolic buffering of Calcium and its influx through voltage-gated Calcium channels. Figure \ref{fig:figuremechanism}A shows a model of a single-compartment model with 8 populations of ion channels (\cite{liuModelNeuronActivityDependent1998}). Without any explicit mechanism for Calcium influx or buffering, the intracellular Calcium levels in this model do not change (Figure \ref{fig:figuremechanism}B) and the model tonically spikes (Figure \ref{fig:figuremechanism}C). Calcium buffering and influx can be modelled by a differential equation that increases intracellular Calcium with the current through Calcium channels and relaxes back to a baseline value (\cite{liuModelNeuronActivityDependent1998, prinzAlternativeHandtuningConductancebased2003, dayanTheoreticalNeuroscience2001}) (Figure \ref{fig:figuremechanism}D). This mechanism exists in the xolotl library as "CalciumMech1" and can be added to the model using a simple "add" statement (Figure \ref{fig:figuremechanism}E). The intracellular Calcium in the model now oscillates periodically (Figure \ref{fig:figuremechanism}F), synchronized to bursts in action potentials in this cell (Figure \ref{fig:figuremechanism}G).

Neurons can regulate their electrical activity by controlling the spectrum of ion channels they express (\cite{macleanActivityIndependentHomeostasisRhythmically2003, turrigianoSelectiveRegulationCurrent1995, schulzVariableChannelExpression2006}). Here, we will show how xolotl can be used to represent a recently proposed model of a homeostatic feedback system that controls the transcription rates of ion channels with the integral of an error signal derived from the intracellular Calcium concentration (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}) (Figure \ref{fig:figuremechanism}H). Since this mechanism affects each ion channel population individually, an object corresponding to this mechanism is added to each "conductance" object in the neuron (Figure \ref{fig:figuremechanism}I). Setting all maximal conductance densities to some low value and integrating the model shows that the intracellular Calcium levels rise over time and approach the target Calcium concentration (Figure \ref{fig:figuremechanism}J), while all conductance densities increase and then remain bounded (Figure \ref{fig:figuremechanism}K). Examining the voltage dynamics of the cell reveal that it transitions from quiescence to truncated bursts of action potentials to periodic bursting as this mechanism regulates the neuron's ion channel spectrum. In summary, our simulator can be used to construct neuron models with intracellular mechanisms such as Calcium influx and buffering, and homeostatic regulation.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		SNAPSHOTS
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Using snapshots to explore model dynamics and parameters}

Switching back and forth points in parameter space and state space of a neuron model is a common occurrence in working with neuron models, and a significant fraction of a modeller's time is spent in a feedback loop of running simulations, viewing the output, changing parameters, and re-running simulations. (\cite{deschutterConsumerGuideNeuronal1992}). xolotl makes it easy to bookmark configurations of a model and return to them at will. Internally, xolotl uses the "serialize" method to gather all parameters and dynamic variables into a vector of values that is passed to the underlying C++ implementation. A paired "deserialize" method is used to update all parameters and variables in the object tree from this vector. This architecture provides a natural framework for representing the state of any model, no matter how complex, using a vector of numbers. The "snapshot" method built into xolotl leverages this schema to save the entire state of the model in a named variable, that can be accessed using another built-in method called "reset".

(Figure \ref{fig:figuresnapshot}) illustrates how these features can be used to explore the dynamics of the model presented in the previous section. First, the current state of the model is saved using the "snapshot" method into a state called ``initial''. In this state, the model exhibits periodic bursting activity due to a particular configuration of maximal conductance densities (Figure \ref{fig:figuremechanism}, orange).  On setting the maximal conductances of the Calcium-permissive channels to 0, the model switches to a tonic spiking activity (Figure \ref{fig:figuremechanism}, purple). Integrating the model for a longer duration allows the homeostatic control mechanism in the cell to restore the conductance profile and bursting activity to a state close to the original state (Figure \ref{fig:figuremechanism}, green). This state is saved using the name ``final''.

The initial state can be returned to using the "reset" method, and a new manipulation to the model can be explored. Here, the intracellular Calcium target is modified, and the model is re-integrated, to yield a different voltage activity (Figure \ref{fig:figuremechanism}, blue). At the end of this numerical exploration, any of the saved states can be quickly returned to using the "reset" method, making the process of re-initializing models to desired states and parameters both error-free and efficient.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		NETWORK MODELS
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Simulating Network Models}

Neurons communicate and interact using synapses, electrochemical junctions between cells (\cite{gjorgjievaComputationalImplicationsBiophysical2016, huaNeuralActivityDynamics2004}). In chemical synapses, the presynaptic neuron releases packets of neurotransmitter across the synaptic cleft, which activate receptors on the postsynaptic neuron. In electrical synapses, no chemical intermediary is involved. Ions flux directly from the presynaptic neuron to the postsynaptic one, setting up a current in the post-synaptic cell. New patterns of activity can emerge from the characteristics of the connecting synapses (\cite{liComodulationSynapsesVoltagegated2018, nadimSynapticDepressionCreates1999, gutierrezRectifyingElectricalSynapses2013, gutierrezMultipleMechanismsSwitch2013}).

Network models in xolotl consist of "compartment" objects that can be connected by "synapse" objects. Two compartments representing different neurons can be connected using synapses using the built-in "connect" method. For example, to connect two single-compartment neurons called "LP" and "PY" using a chemical synapse of type "Cholinergic" with strength 30 nS, we can use

\begin{lstlisting}[style=Matlab-editor]
x.connect('LP','PY','Cholinergic','gbar',30); 
\end{lstlisting}

Our simulator has several types of synapses like Cholinergic, Glutamatergic, AMPA and NMDA built-in, and other synapse classes can be easily added using templates. Figure \ref{fig:figurenetwork} demonstrates the implementation of a model of the triphasic pyloric rhythm in the stomatogastric ganglion of crustaceans (\cite{prinzSimilarNetworkActivity2004}). The pyloric model contains three compartments ("AB", "LP", and "PY") and seven synapses (Figure \ref{fig:figurenetwork}A). This structure is recapitulated in the hierarchy of the xolotl object, where conductances are contained within compartments (Figure \ref{fig:figurenetwork}B). The membrane potentials show triphasic rhythmicity in the three compartments (Figure \ref{fig:figurenetwork}C-E). When the "PY" is depolarized, the dynamical variable mediating the  glutamatergic ("Glut") synapse model between "PY" and "LP" is close to 1 and "LP" is inhibited (Figure \ref{fig:figurenetwork}F). Conversely, when "PY" is hyperpolarized, the dynamical variable is close to 0. When "PY" spikes, IPSPs can be seen in the "LP" voltage trace (Figure \ref{fig:figurenetwork}D-E).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		GRAPHICAL INTERFACE
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Using the GUI to manipulate parameters}

Conductance-based neuron models are typically high dimensional and contain several parameters. Changing a single parameter monotonically can cause non-monotonic changes in behavior of the model, and certain dynamical features may only emerge when in specific non-convex regions of parameter space (\cite{golowasch2002failure}). It is often challenging to build intuition about what effect a parameter has on the model under these conditions. Traditionally, the technique used by computational neuroscientists in building intuition about these models is to iteratively run simulations, view outputs and change parameters. In practice, this meant writing a script, running it, inspecting the output, changing parameters in the script, and repeating this process. It can be cumbersome, and every step in this process involves ``mode'' changes: between a text editor, viewing a graphical output, and the command line that can frustrate the researcher.

Our software is dfesigned to streamline this process and allows for any parameter in any model to be manipulated using graphical sliders, with immediate, real-time feedback of its behavior. Any model in our simulator can be manipulated using

\begin{lstlisting}[style=Matlab-editor]
x.manipulate 
\end{lstlisting}

This method creates a GUI element with sliders for every parameter, and also creates a set of plots that shows the dynamical behavior of the model (Fig. \ref{fig:figuremanipulate}). By default, time series of the voltage and the calcium of every compartment are shown, though this can be modified. Moving any of the sliders updates the value of that parameter in the model, and also triggers a function call that reintegrates the model and updates the output plots. This function call can also update custom plots, like the one shown in (Fig. \ref{fig:figuremanipulate}B). Any model can be manipulated in this way without writing any additional code.

This feature is only possible due to our architectural decision to split the code base across two programming environments. A rich scientific programming environment like MATLAB makes it possible to generate user interface elements and to bind them to data in plots, while the sheer speed of compiled languages like C++ allow for the immediate, real-time feedback and updating of plots. By default, any parameter in the model can be manipulated, including parameters in user-defined mechanisms that do not exist in the base simulator. The GUI can be constrained to arbitrary subsets of model parameters using wild card matching (as shown in Fig. \ref{fig:figuremanipulate}) or by manually specifying the parameters of interest.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		BENCHMARKS
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Benchmarks}
\label{benchmarks}

Our goal in designing xolotl was to create an easy-to-use neuron and network simulator that was fast enough and accurate enough for routine use by computational neuroscientists. In this section, we benchmarked the speed and accuracy of xolotl in simulating single and large numbers of Hodgkin-Huxley-like neuron models (as in Fig. \ref{fig:figurehh}) and bursting neuron models based on the bursting neurons in the lobster stomatogastric ganglion (STG) (\cite{prinzSimilarNetworkActivity2004}). For each type of neuron model, we compared our software to  NEURON (\cite{hinesNEURONSimulationEnvironment1997}), a high-performance and powerful neuron simulator specialized in simulating neurons with complex morphologies and DynaSim (\cite{sherfeyDynaSimMATLABToolbox2018}), a general-purpose simulator that can solve coupled differential equations numerically. All simulators were run on the same hardware using fixed time-step solvers: xolotl used the Exponential Euler method (\cite{dayanTheoreticalNeuroscience2001}), NEURON used the  implicit Euler solver (\cite{hinesNEURONSimulationEnvironment1997}) and DynaSim used C-compiled 2\textsuperscript{nd}-order Runge-Kutta integration scheme as recommended for high-performance (\cite{sherfeyDynaSimMATLABToolbox2018}). We measured the speed of each simulator by dividing the time simulated for, by the time it took for the simulator to complete integration. For example, if a simulator could simulate 10 seconds of model data in 1 second, its speed would be 10X. We measured the speed of every simulator as a function of simulation time step, total length of simulation, and system size.

All three simulators were faster with larger time steps, since fewer iterations were needed (Fig. \ref{fig:figurebenchmark}A-B), and were approximately linear in the region tested. Our simulator compared favorably to NEURON and DynaSim in this task. We also measured the quality of the simulated output by comparing it to the simulated output at the smallest time step. Simulation error was measured using the LeMasson cost (\cite{lemassonIntroductionEquationSolving2000}), and was comparable amongst the three simulators (Fig. \ref{fig:figurebenchmark}C-D). Since our simulator sets up and runs the simulation in C++, it needs to transfer parameters and data to and from the underlying implementation. To measure the performance cost of this overhead, we repeated these benchmarks on all three simulators at a fixed time step of $0.1$ ms and varied the length of time simulated for. Speed increased with simulation duration up to a point, and then saturated, indicating a fixed performance cost to the overhead (Figure \ref{fig:figurebenchmark}E-F, black lines). Simulations using DynaSim, which also used a similar architecture and need to move data between C implementations and the MATLAB workspace, showed a similar increase in speed with simulation length (Figure \ref{fig:figurebenchmark}E-F, red lines). However, simulations using NEURON ran at a constant speed irrespective of simulation length, presumably due to differences in the underlying implementation (Figure \ref{fig:figurebenchmark}E-F, blue lines).

Many simulators have been designed with a focus on simulate large numbers of compartments, either as networks with many identical neurons or in a large multi-compartment neuron model (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003}). While our software is not designed for this task {\em per se}, we measured its performance as a function of the number of compartments simulated. xolotl can quickly create a number of identical copies of a compartment using the "replicate" method:

\begin{lstlisting}[style=Matlab-editor]
x.replicate('compartment_to_replicate',n_copies); 
\end{lstlisting}

We used the "replicate" method to create and run models with varying numbers of neurons (either Hodgkin-Huxley-like or bursting neurons) and measured the speed of all three simulators as a function of system size. Plotting the speed normalized by the system size vs. the system size, we observed that the speed of integration of xolotl is linear with system size (Figure \ref{fig:figurebenchmark}G-H, black lines), for up to 1000 single-compartment neurons (up to 13,000 ODEs). Its performance compares favorably with that of NEURON and DynaSim as a function of system size (Figure \ref{fig:figurebenchmark}G-H).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		DISCUSSION
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Discussion}
\label{discussion}

We set out to design a neuron and network simulator that could be useful in the classroom setting, especially for students of computational science, while also being powerful, fast, and extensible enough to be used for research. By using a novel architecture that permits the symbolic manipulation of C++ objects in a intuitive MATLAB interface, we demonstrated some of the features of our simulator using simulations of single-compartment models of Hogkin-Huxley like neurons (Fig. \ref{fig:figurehh}); voltage clamp experiments to recover activation functions of single channel types (Fig. \ref{fig:figureclamp}); a neuron model where intracellular mechanisms can control the dynamics of Calcium and can regulate the maximal conductance of ion channel types in an activity-dependent manner (Fig. (\ref{fig:figuremechanism}); and a network of neurons with multiple synapse types (Fig. \ref{fig:figurenetwork}). We also illustrated how built-in features of our simulators make it easy to bookmark and jump between model configurations (Fig. \ref{fig:figuremechanism}, purple), and how parameters of the model can be changed using sliders and their effect can be viewed in real time (Fig. \ref{fig:figuremanipulate}).

\subsection{A focus on usability}

``About half the time spent on a typical simulation project involves creating and tuning the model. Thus, a good user interface may contribute more to the overall efficiency of a project than pure computation speed.'' (\cite{deschutterConsumerGuideNeuronal1992}). xolotl is designed primarily with ease-of-use in mind. This includes the time it takes to install, setup, and learn how to use the software, the time to write and debug scripts, and the time to perform the simulations (\cite{rudolph2007much}). An easy-to-use simulation environment must minimize time spent in all these domains, especially during human engagement with the software. Complicated software remains broadly inaccessible and time-consuming even to perform single-compartment simulations, though the actual simulation time may be very small.

We have focussed on making our software as easy to use as possible, without sacrificing performance or extensibility. For example, the software and all dependencies can be installed using a single-line installer script from within the MATLAB command line. The installation includes worked example scripts that demonstrate various features of the simulator, that can be run without any configuration.  We decided to built the the front-end interface to xolotl in MATLAB to facilitate interoperability with existing tools for time series analysis, optimization, and parallel computing. This allowed us to build rich tools for visualization and to interact with the simulation.

\subsection{Comparison with other simulators}

Over the years, several simulators have been developed to integrate systems of coupled differential equations, those that model the spiking activity of neurons (\cite{bretteSimulationNetworksSpiking2007, sherfeyDynaSimMATLABToolbox2018, vitayANNarchyCodeGeneration2015, delormeSpikeNETEventdrivenSimulation2003, hinesNEURONSimulationEnvironment1997}). A critical architectural choice in designing a simulator is how much ``scaffolding'' a user is provided to construct a model; whether a model is specified by equations or by components, or by some combination of the two. In an equation-oriented architecture, the user starts with a blank slate and the primary method of specifying a model is to write out its differential equations. In contrast, in a component-oriented architecture, the primary method of specifying a model is to assemble it from pre-existing components, each of which include differential equations, parameters, and solvers. Both approaches have advantages and disadvantages that are illustrated below.

An equation-oriented simulator can be more transparent and allows the user to know exactly what is being solved, but equations can be cumbersome to write out, read, or to debug. In most commonly used programming environments, these equations have to be entered as strings, and complex parsing has to be carried out by the simulator to check that these strings constitute valid equations. In addition, parameters have to be written explicitly into equations, and it is usually not trivial to change parameters after initial specification. In contrast, components are easy to assemble into a model, but it can be hard to know what they contain, where they are physically located on the user's computer, and how they can be changed. NEURON is primarily a component-oriented neuronal simulator, and new components are specified in special model files that can be ``inserted'' into a model. BRIAN is an equation-oriented neuronal simulator meant to be used from within Python (\cite{goodmanBrianSimulator2009}). XPP is a general purpose dynamical system simulator that is equation-oriented (\cite{ermentrout2002simulating}). DynaSim is an equation-oriented simulator with some component-oriented capabilities. Models can be specified by both strings of equations or components, but since models do not exist as objects in the workspace, parameters and variables have to be reinitialized when changed (\cite{sherfeyDynaSimMATLABToolbox2018}).

Our simulator is a purely component-based simulator, and all equations need to be included in a C++ header file that specifies a object. Since our automatic type system binds MATLAB objects to the underlying C++ header files, objects can be inspected and parameters can be modified in the MATLAB workspace. To mitigate some of the drawbacks of the component-oriented paradigm, we have implemented an architecture that allows the user to access the underlying C++ code of any object by simply clicking on the object tree in the MATLAB command line. This feature removes the uncertainty inherent in other component-oriented simulators of the equations underlying each component, and allows the user to modify these equations if needed.

Another architectural choice in simulators is the syntax required for specifying models. Equation-oriented simulators specify models as strings of equations, so must invent a new syntax to specify derivatives, variables, and other common elements in coupled differential equations. Some component-oriented simulators like NEURON also specify their own syntax, or have invented their own language to specify components and write out equations. While this allows for powerful features such as support for units in NEURON, the user is required to learn a novel syntax and vocabulary, hindering ease of use. In general, the syntax for model specification in different simulator can be different. For example, DynaSim, BRIAN, XPP, and NEURON all use different, incompatible formalisms to represent equations, increasing the cognitive load of users using more than one simulator. In our software, we have elected not to specify a domain-specific ``middleware'' layer, and instead specify and implement models and equations in idiomatic C++. This greatly decreases the learning curve and allows users with a general familiarity with programming languages to quickly acquaint themselves even with the most technical parts of our simulator.

\subsection{Auditability and Reproducibility}

The use of computational tools is increasingly central to the scientific method; yet, the lack of auditability and accountability in their use has led to a crisis of credibility affecting many scientific fields (\cite{stoddenEnhancingReproducibilityComputational2016, bakerWhyScientistsMust2016}). Unlike in experimental research, where a lack of reproducibility can manifest due to meaningful reasons like uncharted differences in experimental protocols or intrinsic variability, the reasons for irreproducibility in computational research are often trivial and include: a) typographical errors from transcribing model parameters and equations, b) obscure software design that leads to users not knowing precisely what equations the software is solving, or what parameters it is using, c) incompatibility with version control systems and rolling software development that leads to ambiguity in which the version of the software that was used to generate a particular result is unclear, and d) convoluted architectures that make it too complex for non-experts to understand the inner workings of the software (\cite{xuMeasurementSourceCode2017, sedanoCodeReadabilityTesting2016, vikstromStudyAutomaticTranslation2009}).

Our software was designed with this threat model in mind, and has features that allow the user to answer the following questions in the affirmative: ``Can I be sure that I am doing what I think I am doing?'' and ``Is it possible for others to reproduce {\em exactly} what I have just done?''. Our goal was to design software that would allow the user to verify for herself that the software was running as she intended, and to be able to reproduce results from others quickly and unambiguously. The primary design choice in our software that enables auditability and reproducibility is that every simulation is tied to an alphanumeric checksum, or hash, using the "MD5" algorithm (\cite{rivestMD5MessageDigestAlgorithm1992}). The hash is computed from every C++ file that is part of the model, and any changes in any C++ file included in the model will trigger a recompilation of that model. Thus, the hash guarantees that a given model is derived from a set of source files, obviating any ambiguity about the code used to generate a model. In addition, every parameter and state variable in the model can also be hashed together with underlying code, allowing the user to generate a short checksum that guarantees with high probability that every aspect of the model -- code, parameters, and initial conditions -- are exactly as they should be.

Typographical errors from transcribing model parameters and equations can also be detected using hashes, and the component-oriented architecture of our software makes it easy to debug code and spot errors. Our software has been designed so that it is possible to explore the model interactively in the command line, and it is possible to ``click through'' from the highest level of the model in the command line all the way down to the underlying code of any component in the model. This design allows the user to know precisely the equations being solved, and view the code that numerically integrates them. Finally, the core of our software is written in a few hundred lines of code and contains just four classes: compartments, conductances, synapses, and mechanisms. This architectural simplicity lets a motivated user understand the entirety of our code quickly.

\subsection{Outlook and future directions}

In its current form, xolotl is an efficient and easy to use neuron and network simulator that is actively being used in research. 
Results from an early version of this simulator guided intuition in the modeling of a recently characterized Calcium-dependent Potassium channel found in {\em Drosophila} neuromuscular presynaptic terminals (\cite{bronk2018regulation}). Work on the simulator continues in the open at a publicly accessible repository (https://github.com/sg-s/xolotl/), and the library of conductances, synapses and mechanisms that xolotl ships with grows continuously.

Since xolotl models are bonafide MATLAB objects, they are compatible with most of the powerful tools that exist within MATLAB. For example, it is possible to write simple scripts that run xolotl models in parallel using MATLAB's parallel processing toolbox, speeding up large simulations. xolotl models are also compatible with the global optimization toolbox in MATLAB, allowing parameters in xolotl models to be optimized, enabling the creation of toolboxes that efficiently tune parameters in neuron models to satisfy arbitrary constraints (\cite{Achard:2006ia, Krichmar:2014jj, Keren:2005ja, VanGeit:2007ea, Druckmann:2008kg}).

Care has been taken to reduce the amount of technical debt (\cite{suryanarayana2014refactoring}) associated with this project, with all parts of the simulator and dependencies written in a modular, objected oriented fashion. As a result, many of the key features and architectures of xolotl can be reused by others in their own applications. For example, cpplab, the automatic type system that binds C++ code to MATLAB objects, is independent of this simulator, and exists as a distinct repository that has been made freely available (https://github.com/sg-s/cpplab); and the framework for generating a GUI with sliders for each parameter that is hooked up to function callbacks also exists as an independent repository (https://github.com/sg-s/puppeteer) that can be easily integrated into other applications.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		PUBLICATION DETAILS
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Conflict of Interest Statement}
%All financial, commercial or other relationships that might be perceived by the academic community as representing a potential conflict of interest must be disclosed. If no such relationship exists, authors will be asked to confirm the following statement:

The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.

\section*{Author Contributions}

SG-S \& AH wrote the software and created the online user documentation. EM supervised the project. All authors wrote and reviewed the paper.

\section*{Funding}

AH received funding from the National Institute on Drug Abuse (NIDA) through the undergraduate training grant in computational neuroscience (1R90DA033463-01).

\section*{Acknowledgments}
The authors would like to thank Timothy O'Leary and Leandro Alonso for useful discussions, and Janis Li for help transcribing many conductance models.

\section*{Data Availability Statement}

The code to generate all figures is available at (\url{https://github.com/marderlab/xolotl-paper}). xolotl is freely available at (\url{https://github.com/sg-s/xolotl}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		BIBLIOGRAPHY
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\bibliographystyle{frontiersinSCNS&ENG} % for Science and Engineering articles
%\bibliography{Bibliography}   % name your BibTeX data base
\printbibliography

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		FIGURES
%   (because it's 2018 and we still have this dumb system where all the figures are at the end. why? who knows
%   why can't they be in a more sensible place? it's not like we're not typesetting this ourselves. why do people
%   go out of their way to make papers unreadable? who actually likes the figures and the captions at the end,
%   separated into their own pages?)
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\section*{Figure captions}

\FloatBarrier

\newpage

\begin{figure}
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_HH}
	\caption{Simulating a single-compartment Hodgkin-Huxley spiking neuron model. (A) Schematic representation of a single-compartment neuron with three populations of ion channels (colored rectangles). (B) In our simulator, the soma is represented using an object of class ``compartment'' and populations of ion channels are represented by ``conductance'' objects contained within the compartment object. (C) The code snippet shown sets up this neuron model, injects current, integrates and plots the voltage, and displays activation functions, all in a few lines of code. (D) Simulated voltage trace of a Hodgkin-Huxley model with three conductances and 0.2 nA of injected current. Colors indicate the dominant current (gold is fast sodium (NaV), blue is delayed rectifier (Kd), red is Leak). (E) firing rate $vs$. current (f-I) curve of this neuron. (F-G) Steady-state gating functions for activation (m) and inactivation (h) gating variables. (H-I) Voltage-dependence of time constants for activation (m) and inactivation (h) gating variables}
	\label{fig:figurehh}
\end{figure}

\begin{figure}[!htb]
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_clamp}
	\caption{Simulating a voltage-clamp experiment. The diagram shows a cell with delayed rectifier potassium conductance (from \cite{liuModelNeuronActivityDependent1998}) that is being recorded from in two-electrode voltage clamp (A). The code snippet shown here sets up a model with a single compartment and a single channel type, and clamps the cell to a constant voltage and integrates it (B). Voltage steps that the cell is clamped to (C). Clamping currents as a function of time (D). Asymptotic clamped current {\em vs.} clamped voltage for this cell (E). Accounting for the reversal potential of Potassium ions yields the conductance-voltage curve of this channel type (F). Normalized conductance-voltage curves, with sigmoid fits with various exponents (G). An exponent of $n=4$ yields the best fit, allowing for the characterization of the activation function of this channel type (H). }
	\label{fig:figureclamp}
\end{figure}

\begin{figure}[!htb]
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_mechanism}
	\caption{Modeling intracellular mechanisms. A single-compartment neuron model with 8 channel types (A). Since there is no mechanism for changing intracellular Calcium in this model, the Calcium level stays constant (B), and the cell tonically spikes (C). Intracellular Calcium buffering and influx through voltage gated Calcium channels (VGCCs) can be modeled using a simple differential equation (D). Code snippet shows how this mechanism can be added to the neuron model (E). The cell now bursts periodically, with synchronized oscillations in intracellular Calcium (F-G). Schematic of Calcium-dependent integral control homeostasis (\cite{olearyCorrelationsIonChannel2013, olearyCellTypesNetwork2014}) (H). In this feedback system, the rates of mRNA synthesis depend on the Calcium level in the cell, which depends on the membrane voltage, which in turn depends on the conductance density of all channel types, which, through translation, depends on the mRNA abundance. (I) The code snippet shows how these integral controllers are implemented as \texttt{mechanism} objects, and can be added to conductances. (J) On integrating the model, intracellular calcium levels rise and approach the target (red dashed line). This is accompanied by an increase in the conductance densities of all channels being controlled by this homeostatic mechanism (K). The voltage behavior of the cell changes from silence to bursting with truncated spikes to regular bursting.}
	\label{fig:figuremechanism}
\end{figure}

\begin{figure}[!htb]
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_snapshot}
	\caption{Snapshots allow the user to bookmark points in parameter and state space of the model and return to them $ad$   $arbitrium$. The initial state (orange node) of the single compartment model in the previous example is saved using the \texttt{snapshot} method. This method saves all parameters and dynamic variables of the model in a named state. The first column (orange plots) shows the profile of conductances and the voltage dynamics of the model at this point at this time. The maximal conductances of the Calcium channels are then set to zero (purple node), changing the voltage dynamics of the neuron (purple plots). After evolving the model for some time (green node), the conductance profile and voltage dynamics returns to a state similar to the initial state (green plots). This configuration is now saved in a state called $final$ and the initial configuration is returned to using the \texttt{reset} method (backwards arrow from green to orange). Another parameter is now changed (the Calcium target), and the model is integrated to reach a new state (blue node) where the voltage dynamics are different from the $initial$ state. In summary, any state can be bookmarked using a descriptive name using the \texttt{snapshot} method, and can be returned to using the \texttt{reset} method.}
	\label{fig:figuresnapshot}
\end{figure}

\begin{figure}[!htb]
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_network}
	\caption{Simulating a network of conductance-based model neurons, coupled by voltage-dependent dynamic chemical synapses.  A three-compartment model of the pyloric network in the crustacean stomatogastric ganglion (\cite{prinzSimilarNetworkActivity2004}) (A). Each neuron is modeled with a single compartment with 7-8 intrinsic conductances and 1-3 post-synaptic conductances. Synapses can be one of two types (Cholinergic (dashed lines) or Glutamatergic (solid lines)) and have different kinetics.  The code snippet shows how neurons can be created, wired together using synapses, and how the model can be integrated to return voltages and intracellular calcium levels in every compartment, and the state of every synapse (B). Simulated voltage trace of a model network for the three compartments obtained from this simulation (C-E) . Time series activation variable of the Glutamatergic synapse between PY and LP (red connection in diagram) shows how the synapse becomes active every time the PY neuron spikes (F).}
	\label{fig:figurenetwork}
\end{figure}

\clearpage

\begin{figure}
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_manipulate}
	\caption{Manipulating neuron parameters in real time. Any set of parameter in the model can be manipulated; here, the maximal conductance of every conductance type in the model from Fig. \ref{fig:figurehh} is being manipulated using the code snippet shown here. The screenshot shows a GUI with sliders for every parameter of interest that is created by the \texttt{manipulate} method. These sliders can be linked to an arbitrary number of visualization functions. In this example, two visualization functions are used: the built in \texttt{plot} method (A) and a custom function that computes the firing-rate-$vs$.-injected current curve for this neuron (B). Both plots refresh themselves with every movement of any slider, allowing the user to build intuition about how every parameter controls the dynamical behavior of the model. A screen recording of this model being manipulated in real time is included in Supplementary Material}
	\label{fig:figuremanipulate}
\end{figure}

\clearpage

\begin{figure}
	\centering
	\includegraphics[width=1.0\linewidth]{gfx/figure_benchmark}
	\caption{Comparison of speed and accuracy of xolotl, NEURON and DynaSim. The top row shows simulations of a tonically-firing Hodgkin-Huxley model with three conductances with constant injected current (as in Fig. \ref{fig:figurehh}). The bottom row shows simulations of  a bursting stomatogastric ganglion neuron model with 8 conductances (as in Fig. \ref{fig:figurenetwork}). Ratio of run-time to simulation time (relative speed) as a function of simulation time step (A, B). Simulation error as a function of the step size (C, D).  Relative speed of integration as a function of the simulation length (E, F). Relative speed of integration, normalised by system size, as a function of the number of compartments simulated simultaneously (G, H). All benchmarks were performed on the same computer, and all simulators using fixed time-step integration methods. NEURON was run using the Python wrapper. }
	\label{fig:figurebenchmark}
\end{figure}

\end{document}